Community-Centric Innovation

### 2014-03-01 paper.tex
working on the conclusion

There is a large number of community networks and they are very heterogeneous.
They use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same layer-2 technology, it is still likely that they use different hardware, firmware or protocols.
The sizes of community networks also vary greatly and can range from few tenths of nodes to tenths of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

@@

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency.
The combination of cryptography, free libre software and a data network makes it possible the use of networked cryptocurrencies.
In these technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules (i.e., a protocol) to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their computational work and the majority decides which transactions are included in the public distributed ledger.

@@

@@

\subsection{Confine}
The Confine project emphasizes the links between research and community networks.
The goal is to make it easier for researchers to run experiments in community networks.
It provides the tools for automatize and monitor the experiments, as well as the hardware distributed across the community network.

@@

\section{Related Work}
\section{Conclusions}

>>>>

There is a large number of community networks and they are very heterogeneous \cite{avonts2013qbe}.
They use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same layer-2 technology, it is still likely that they use different hardware, firmware or protocols.
The sizes of community networks also vary greatly and can range from few tenths of nodes to tenths of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

@@

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency \cite{nakamoto2008bpe}.
The combination of cryptography, free libre software and a data network makes it possible the use of networked cryptocurrencies.
In these technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules (i.e., a protocol) to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their computational work and the majority decides which transactions are included in the public distributed ledger.

@@

Decentralization, user empowerment and a focus on community-centric models seem to be a common denominator in the most exciting innovations in the area of information and communication technologies.

@@

\subsection{Confine}
The Confine project \cite{braem2013crc} emphasizes the links between research and community networks.
The goal is to make it easier for researchers to run experiments in community networks.
It provides the tools for automatize and monitor the experiments, as well as the hardware distributed across the community network.

@@

\subsection{D-Cent}
The area of cryptocurrency is covered by the D-Cent project.
This project is actually more ambitious and also considers decentralized social networks and tools for a more direct engagement in social decision making and deliberation.

The scope of this project also includes community-centric innovation, privacy and security, open knowledge, open data and open standards.

\section{Conclusions}
In some aspects of information and communication technologies, communities are  leading the path of innovation.
Community-centric innovation is characterized by peer-to-peer relations and openness.
Communities have also demonstrated their ability to take advantage of data networks and network effects.
In the case of community networks, a reinforcement loop is established in which a community deploys a network that, in its turn, strengthens the community.
Community networks have been with us for over a decade and have remained faithful to the original spirit of the Internet in terms of decentralization.
These networks are planned, financed, deployed and maintained by the same people that uses them.
They adopt self-management dynamics and can benefit from the combined knowledge of all the members of the community.

Despite being more time-consuming and less reliable for the user, they are still a valid alternative for some people.
The possibility of controlling the network and being participants instead of consumers are some of the reasons for joining a community network.
Another reason is that the idea of freedom is embedded in community networks.
Community networks can resist authoritarian censorship or Internet blackouts ordered by governments.
They are also extremely useful in case of a natural disaster when centralized infrastructures may collapse.

Networked cryptocurrencies are another example of community-centric innovation.
These currencies are developed by the communities as a result of open debate and discussion.
There are many similitudes with community networks.
The users of networked cryptocurrencies can contribute one node to the network to help in maintaining security.
The members of the community can also participate in many other ways offering products and services and using the cryptocurrency in their transactions.Networked cryptocurrencies
